Replication Script to JSS Submission 1518:\\ \textsl{Flexible Simulation of Event Histories: A~Guide for the \textsf{R}~Add-on~Package~\texttt{simMSM}}

### 2016-03-09 JSS/jss-style/article.tex
meeting update

We now convert it into an \pkg{igraph} object (\citealt{igraph}) \code{ig} using the function \code{dfToIG()}.

\begin{CodeChunk}
\begin{CodeInput}
R> ig <- dfToIG(sbGeneal)
R> ig
\end{CodeInput}
\begin{CodeOutput}
IGRAPH UNW- 230 340 -- 
+ attr: name (v/c), weight (e/n)
+ edges (vertex names):
 [1] 5601T    --Hutcheson        Adams    --Dunfield        
 [3] A.K.     --A.K. (Harrow)    Altona   --Flambeau        
 [5] Amcor    --Amsoy 71         Adams    --Amsoy           
 [7] Amsoy 71 --C1253            Anderson --Lincoln         
 [9] Bay      --York             Bedford  --Forrest         
[11] Beeson   --Kent             Blackhawk--Richland        
[13] Bonus    --C1266R           Bradley  --J74-39          
[15] Bragg    --Jackson          Bragg    --Bragg x D60-7965
+ ... omitted several edges
\end{CodeOutput}
\end{CodeChunk}

There are many statistics about the \code{sbGeneal} genealogical dataset that we may wish to know that cannot easily be obtained through images and tables. The package function \code{getBasicStatistics()} can be called, using the \code{ig} object as input. This will return a list of common graph theoretical measurements regarding the genealogical graph structure. For instance, is the whole structure connected? If not, how many separated components does it contain? In addition to these statistics, the \code{getBasicStatistics()} function will also return the number of nodes, the number of edges, the average path length, the graph diameter, and other graph theoretical information.

\begin{CodeChunk}
\begin{CodeInput}
R> getBasicStatistics(ig)
\end{CodeInput}
\begin{CodeOutput}
$isConnected
[1] FALSE

@@

\begin{CodeChunk}
\begin{CodeInput}
R> pathTN <- getPath("Tokyo", "Narow", ig, sbGeneal)
R> pathTN
\end{CodeInput}
\begin{CodeOutput}
$pathVertices
[1] "Tokyo"    "Volstate" "Jackson"  "R66-873"  "Narow"

@@

\begin{Code}
R> pathZB <- getPath("Zane", "Bedford", ig, sbGeneal)
R> plotPath(pathZB)
\end{Code}

@@

\begin{Code}
R> plotAllImage <- plotPathOnAll(pathTN, sbGeneal, ig, binVector = 1:3,
  pathColor = "red")
R> plotAllImage + ggplot2::theme(axis.text = element_text(size = 20),
  axis.title = element_text(size = 20))
R> plotAllImage <- plotPathOnAll(path = path, geneal = sbGeneal, ig = ig,
  binVector = sample(1:3), edgeCol = "gray84", pathEdgeCol = "blue", nodeSize
  = 1, pathNodeSize = 3)
\end{Code}

\begin{figure}%[h]
    \centering
    \includegraphics[width=\textwidth]{fig3}
    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 3.}
    \label{fig:fig3}
\end{figure}

@@

Hence, in the code above, the user selected a \code{binVector} value of three, and a \code{plotColor} of red, which produces the plot in Figure \ref{fig:fig3}. We see that edges not on the path of interest are thin and gray by default, whereas edges on the path of interest are bolded and red. We also see that varieties in the path of interest are boldfaced by default.

@@

\begin{Code}
R> plotAllImage <- plotPathOnAll(pathTN, sbGeneal, ig, binVector = 1:6,
  pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
\end{Code}

\begin{figure}%[h]
    \centering
    \includegraphics[width=\textwidth]{fig4}
    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
    \label{fig:fig4}
\end{figure}

This similar code now outputs the plot shown in Figure \ref{fig:fig4}. We can immediately see that this plot more successfully mitigates text variety label overlap than the previous plot in \ref{fig:fig3}. We can also confirm what we saw in the previous plot that indeed most varieties have development years between 1950 and 1970, and any textual overlap is confined to this range of years.

@@

As such, we plan to add a feature to the \pkg{ggenealogy} package that allows users to manually fine-tune the plot they deem best after examining various bin sizes. For example, after comparing several bin sizes (1-12) on the current soy bean data set, we determined that the bin size of 6 produced minimal textual overlap, as seen in Figure 4.

@@

\begin{Code}
>R varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow",
  "Pella", "Tokyo", "Young", "Zane")
>R plotDegMatrix(varieties, ig, sbGeneal, "Variety", "Variety", "Degree") +
  ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 15),
  legend.text = ggplot2::element_text(size = 15))
\end{Code}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{degMatrix}
    \caption{The shortest path degree matrix between ten varieties of interest.}
    \label{fig:degMatrix}
\end{figure}

@@

\section{Academic Genealogy of Statisticians Dataset}

In addition to the soybean genealogy, the \proglang{ggenealogy} package also comes with an academic genealogy of statisticians. To develop this later dataset, we contacted the Mathematics Genealogy Project, a web-based database for the genealogy of academic mathematicians. This database, which currently contains almost 200,000 entries, is a service of the North Dakota State University Department of Mathematics and the American Mathematical Society. The Mathematics Genealogy Project contact provided us a Structured Query Language (SQL) export, and we used PostgreSQL to query the database (\citealt{psql}, \citealt{sql}).

Each entry in the database contained 26 variables pertaining to an individual who received a graduate-level academic degree in statistics. One of these variables was called ``msc" (Mathematics Subject Classification), and we selected only those entries that contained a value of 62 for this variable (coded as ``Statistics"). This process resulted in 8995 entries, which we further reduced by only retaining those entries that contained an advisor who was also in the field ``Statistics". For each of the remaining 3293 entries, we only maintained 6 of the variables. This process resulted in the second example dataset provided in the \proglang{ggenealogy} package

This later example dataset is available in a data frame structure that contains 3293 rows, each which corresponds to a parent-child relationship, where the child is an individual who received a graduate-level academic degree in statistics. The parent is the advisor

\begin{CodeChunk}
\begin{CodeInput}
R> data(statGeneal)
R> dim(statGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] 3293    6
\end{CodeOutput}
\begin{CodeInput}
R> colnames(statGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] "child"   "parent"  "year"    "country" "school"  "thesis"
\end{CodeOutput}
\end{CodeChunk}

which is available in a data frame structure with 412 rows, each representing a direct parent-child relationship between a pair of soybean varieties. In total, there are 230 unique soybean varieties in the dataset. These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s. They also contain information on the developmental years, as well as the copy number variants, single nucleotide polymorphisms, protein content, and yield, of each of the soybeans.

@@

\begin{figure}%[h]
    \centering
    \includegraphics[width=\textwidth]{dCox.png}
    \caption{plot7}
    \label{fig:dCox}
\end{figure}

>>>>

We now convert it into an \pkg{igraph} object (\citealt{igraph}) \code{sbIG} using the function \code{dfToIG()}.

\begin{CodeChunk}
\begin{CodeInput}
R> sbIG <- dfToIG(sbGeneal)
R> sbIG
\end{CodeInput}
\begin{CodeOutput}
IGRAPH UNW- 230 340 -- 
+ attr: name (v/c), weight (e/n)
+ edges (vertex names):
 [1] 5601T    --Hutcheson        Adams    --Dunfield        
 [3] A.K.     --A.K. (Harrow)    Altona   --Flambeau        
 [5] Amcor    --Amsoy 71         Adams    --Amsoy           
 [7] Amsoy 71 --C1253            Anderson --Lincoln         
 [9] Bay      --York             Bedford  --Forrest         
[11] Beeson   --Kent             Blackhawk--Richland        
[13] Bonus    --C1266R           Bradley  --J74-39          
[15] Bragg    --Jackson          Bragg    --Bragg x D60-7965
+ ... omitted several edges
\end{CodeOutput}
\end{CodeChunk}

There are many statistics about the \code{sbGeneal} genealogical dataset that we may wish to know that cannot easily be obtained through images and tables. The package function \code{getBasicStatistics()} can be called, using the \code{sbIG} object as input. This will return a list of common graph theoretical measurements regarding the genealogical graph structure. For instance, is the whole structure connected? If not, how many separated components does it contain? In addition to these statistics, the \code{getBasicStatistics()} function will also return the number of nodes, the number of edges, the average path length, the graph diameter, and other graph theoretical information.

\begin{CodeChunk}
\begin{CodeInput}
R> getBasicStatistics(sbIG)
\end{CodeInput}
\begin{CodeOutput}
$isConnected
[1] FALSE

@@

\begin{CodeChunk}
\begin{CodeInput}
R> pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal)
R> pathTN
\end{CodeInput}
\begin{CodeOutput}
$pathVertices
[1] "Tokyo"    "Volstate" "Jackson"  "R66-873"  "Narow"

@@

\begin{Code}
R> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
R> plotPath(pathBZ)
\end{Code}

@@

\begin{Code}
R> plotAllImage <- plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3,
  pathColor = "red")
R> plotAllImage ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
R> plotAllImage <- plotPathOnAll(path = path, geneal = sbGeneal, ig = sbIG
  binVector = sample(1:3), edgeCol = "gray84", pathEdgeCol = "blue", nodeSize
  = 1, pathNodeSize = 3)
\end{Code}

\begin{figure}%[h]
    \centering
    \includegraphics[width=\textwidth]{plotTNBin3}
    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 3.}
    \label{fig:plotTNBin3}
\end{figure}

@@

Hence, in the code above, the user selected a \code{binVector} value of three, and a \code{plotColor} of red, which produces the plot in Figure \ref{fig:plotTNBin3}. We see that edges not on the path of interest are thin and gray by default, whereas edges on the path of interest are bolded and red. We also see that varieties in the path of interest are boldfaced by default.

@@

\begin{Code}
R> plotAllImage <- plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6,
  pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
\end{Code}

\begin{figure}%[h]
    \centering
    \includegraphics[width=\textwidth]{plotTNBin6}
    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
    \label{fig:plotTNBin6}
\end{figure}

This similar code now outputs the plot shown in Figure \ref{fig:plotTNBin6}. We can immediately see that this plot more successfully mitigates text variety label overlap than the previous plot in \ref{fig:plotTNBin3}. We can also confirm what we saw in the previous plot that indeed most varieties have development years between 1950 and 1970, and any textual overlap is confined to this range of years.

@@

As such, we plan to add a feature to the \pkg{ggenealogy} package that allows users to manually fine-tune the plot they deem best after examining various bin sizes. For example, after comparing several bin sizes (1-12) on the current soy bean data set, we determined that the bin size of 6 produced minimal textual overlap, as seen in \ref{fig:plotTNBin6}.

@@

\begin{Code}
>R varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow",
  "Pella", "Tokyo", "Young", "Zane")
>R plotDegMatrix(varieties, sbIG, sbGeneal, "Variety", "Variety", "Degree") +
  ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 15),
  legend.text = ggplot2::element_text(size = 15))
\end{Code}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{degMatrix}
    \caption{The shortest path degree matrix between ten varieties of interest.}
    \label{fig:degMatrix}
\end{figure}

@@

\section{Academic Genealogy of Statisticians}

In addition to the soybean genealogy, the \proglang{ggenealogy} package also comes with an academic genealogy of statisticians. To develop this later dataset, we contacted the Mathematics Genealogy Project, a web-based database for the genealogy of academic mathematicians. This database, which currently contains almost 200,000 entries, is a service of the North Dakota State University Department of Mathematics and the American Mathematical Society. The Mathematics Genealogy Project contact provided us a Structured Query Language (SQL) export, and we used PostgreSQL to query the database (\citealt{psql}).

Each entry in the database contained 26 variables pertaining to an individual who received a graduate-level academic degree in statistics. One of these variables was called ``msc" (Mathematics Subject Classification), and we selected only those entries that contained a value of 62 for this variable (coded as ``Statistics"). Furthermore, we only retained those entries that, if they contained an advisor, then that advisor was also in the field ``Statistics". This process resulted in 8995 entries, which we reduced to 8166 entries by removing duplicate entries. With the final data frame of 8995 entries, we only maintained 6 of the variables.

This process resulted in the second example dataset provided in the \proglang{ggenealogy} package. This dataset is structured as a data frame that contains 8166 rows, each of which corresponds to a parent-child relationship, where both the child and the parent received postbaccalaureate degrees in statistics, and the parent was the academic advisor to the child. The columns of the dataset provide information about the year the child obtained the degree, the country and school from which the child obtained the degree, and the thesis title of the degree awarded to the child.

\begin{CodeChunk}
\begin{CodeInput}
R> data(statGeneal)
R> dim(statGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] 8166    6
\end{CodeOutput}
\begin{CodeInput}
R> colnames(statGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] "child"   "parent"  "year"    "country" "school"  "thesis"
\end{CodeOutput}
\end{CodeChunk}

%\section{Plotting Ancestors and Descendants by Generation}

This second example dataset of academic genealogy is of a slightly different structure than the first example dataset of agronomical genealogy, because each child no longer requires exactly two parents. The ability to plot ancestors and descendants by generation was demonstrated on the agronomical genealogy in Figure \ref{fig:Lee}. As we believe this is the most novel plotting tool in the \proglang{ggenealogy} package, we can test it again here using the academic genealogy.

We can determine which academic statistician in this \code{statGeneal} dataset has the largest number of descendants. To do so, we create a vector \code{indVec} that contains the names of all individuals (whether child or parent) in the dataset. We then use the \proglang{dplyr} package to apply the ggenealogy function \code{getDescendants} on each individual in the \code{indVec} vector (\citealt{dplyr}). We set the parameter \code{gen} to a conservatively large value of 100 as this dataset is unlikely to have any individuals with more than 100 generations of descendants.

We then create a table to examine all values of descendant counts observed in the dataset, along with the number of individuals who had each of those values of descendant counts. Of the 8166 individuals in this dataset, 6252 of them have zero descendants, 322 of them have one descendant, and 145 of them have two descendants. There are only 17 individuals who have more than 30 descendants, and there is one individual who has the largest value of 159 descendants. We determine that this individual is the prominent British statistician Sir David Cox, who is known for the Box-Cox transformation and Cox processes, as well as for mentoring many younger researchers who later became notable statisticians themselves.

\begin{CodeChunk}
\begin{CodeInput}
>R library(dplyr)
>R indVec <- unique(c(statGeneal$child, statGeneal$parent))
>R indVec <- indVec[which(indVec != "", )]
>R dFunc <- function(var) nrow(getDescendants(var, statGeneal, gen = 100))
>R numDesc <- sapply(indVec, dFunc)
>R table(numDesc)
\end{CodeInput}
\begin{CodeOutput}
numDesc
   0    1    2    3    4    5    6    7    8    9   10   11   12   13   14 
6252  322  145   88   58   36   31   22   23   14   17   13   14    9    9 
  15   16   17   18   19   20   21   22   23   24   25   26   27   29   30 
   6    4    3    2    5    7    5    3    3    2    2    6    1    1    3 
  34   37   38   40   41   44   45   48   49   61   62   75   77   84  159 
   2    1    1    1    1    1    1    1    2    1    1    1    1    1    1
\end{CodeOutput}
\begin{CodeInput}
R> which(numDesc == 159)
\end{CodeInput}
\begin{CodeOutput}
David Cox 
     1980
\end{CodeOutput}
\end{CodeChunk}

We can now visualize how these 159 descedants are related to Sir David Cox by plotting parent-child relationships, with generation count constricted on the horizontal axis, similar to what we performed in Figure \ref{fig:Lee}. We create this in Figure \ref{fig:dCox} using the code below.

\begin{CodeInput}
R> plotAncDes("David Cox", statGeneal, mAnc = 6, mDes = 6, vCol = "blue")
\end{CodeInput}

\begin{figure}%[h]
    \centering
    \includegraphics[width=\textwidth]{dCox.png}
    \caption{The 159 academic descendants of Sir David Cox.}
    \label{fig:dCox}
\end{figure}

We see from Figure \ref{fig:dCox} that Sir David Cox had 42 children, many of them becoming notable statisticians, such as Basilio Pereira, Valerie Isham, Gauss Cordeiro, Peter McCullagh, and Henry Wynn. The child of his who produced the most children of his own was Peter Bloomfield, who has 26 children and 49 descendants. In total, Sir David Cox has five generations of statistics students in this dataset.

\begin{CodeChunk}
\begin{CodeInput}
R> length(getChild("Peter Bloomfield", geneal))
\end{CodeInput}
\begin{CodeOutput}
[1] 26
\end{CodeOutput}
\begin{CodeInput}
R> nrow(getDescendants("Peter Bloomfield", geneal, gen = 100))
\end{CodeInput}
\begin{CodeOutput}
[1] 49
\end{CodeOutput}
\end{CodeChunk}

It would be insightful to examine a more detailed view of one of the longest strings of parent-child relationships between Sir David Cox and one of his fifth generation descendants.

\begin{CodeChunk}
\begin{CodeInput}
R> statIG <- dfToIG(statGeneal)
R> pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal, isDirected
  = FALSE)
R> plotPath(pathCB) + ggplot2::theme(axis.text = ggplot2::element_text(size
  = 12), axis.title = ggplot2::element_text(size = 12)) +
  ggplot2::scale_x_continuous(breaks = seq(1940, 2010, by = 10))
\end{CodeInput}
\end{CodeChunk}

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{pathCB}
    \caption{The shortest path between Sir David Cox and Petra Buzkova is strictly composed of five unidirectional parent-child relationships that span about 55 years. We see that the time difference between when an advisor and student earned their degrees is not consistent across this path: The three statisticians who earned their degrees earliest in this path span more than about 30 years in degree aquisition, whereas the three statisticians who earned their degrees later in this path only span less than about 10 years in degree acquisition.}
    \label{fig:pathCB}
\end{figure}

\begin{figure}%[h]
    \centering
    \includegraphics[width=\textwidth]{plotCBText}
    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
    \label{fig:plotCBText}
\end{figure}

\begin{figure}%[h]
    \centering
    \includegraphics[width=\textwidth]{plotCBNoText}
    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
    \label{fig:plotCBNoText}
\end{figure}

@@

### 2016-03-09 JSS/jss-style/replicationScript.tex
meeting update

\documentclass{article}
\parindent 0pt
\title{Replication Script to JSS Submission:\\
\textsl{ggenealogy: An R Package for Visualizing Genealogical Data}
\author{Lindsay Rutter, Susan VanderPlas, Di Cook}
\date{\today}
\begin{document}
\maketitle
\SweaveOpts{concordance=TRUE}
\tableofcontents
<<>>=
install.packages("ggenealogy")
library("ggenealogy")
@
%%%%%%%%%%%%%%%
%% Section 1 %%
%%%%%%%%%%%%%%%
\section{Generating a Graphical Object}

Examine structure of sbGeneal example dataset:

<<>>=
data(sbGeneal)
str(sbGeneal)
@

Convert sbGeneal into igraph object:

<<>>=
library("mstate")
tra2 <- !is.na(transMat(list(c(2), c(1))))
tra2
@
Use \texttt{mplskeleton} to construct a skeleton of the model parameter list \texttt{mpl}:
<<>>=
mpl <- mplskeleton(tmat = tra2)
str(mpl)
@
Parametrize the model and complete \texttt{mpl}:
<<>>=
mpl[[1]]$bhr[[2]] <- mpl[[2]]$bhr[[1]] <- function(t){return(0.5)}
## time-varying x2 and time-varying f(x2):
mpl[[1]]$eta[[2]] <- function(x.i, t){
  ifelse(t < 5, 
         return(1.0 * x.i[1] + 0.5 * x.i[2]), 
         return(1.0 * x.i[1] + 1.0 * x.i[3]))}
## time-varying x2 and time-varying f(x1):
mpl[[2]]$eta[[1]] <- function(x.i, t){
  ifelse(t < 5, 
         return(-0.5 * x.i[1] + 0.5 * x.i[2]), 
         return( 1.0 * x.i[1] + 0.5 * x.i[3]))}
@
Simulate event histories:
<<>>=
set.seed(123)
N <- 500
X <- matrix(nrow = N, ncol = 2, rnorm(2 * N))
X <- cbind(X, X[, 2] + runif(N)/10)
colnames(X) <- c("x1", "x2.t1", "x2.t2")
head(X)
Xstruc <- matrix(ncol = 2, nrow = 2, data = 0)
rownames(Xstruc) <- c("t1", "t2")
colnames(Xstruc) <- c("x1", "x2")
Xstruc[, 1] <- 1
Xstruc[, 2] <- c(2, 3)
Xstruc
d <- simeventhistories(n = N, mpl = mpl, X = X, max.time = 10, 
                       change.times = c(5), Xstruc = Xstruc)
head(d$msm.basics)
head(d$ttsce[, 1:7])
head(d$tt.indicators)
@
Nelson-Aalen estimator of the cumulative hazard rate functions:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
plotnae(d = d$msm.basics, mpl = mpl, xlim = c(0, 15))
abline(a = 0, b = 0.5, col = "darkgrey")
@

>>>>

\documentclass{article}
\parindent 0pt
\title{Replication Script to JSS Submission 1518:\\
\textsl{Flexible Simulation of Event Histories: A~Guide for the \textsf{R}~Add-on~Package~\texttt{simMSM}}}
\author{Holger Reulen\\ University of G\"ottingen}
\date{\today}
\begin{document}
\maketitle
\SweaveOpts{concordance=TRUE}
\tableofcontents
<<>>=
options(prompt = " ", continue = "  ")
install.packages("simMSM")
library("simMSM")
@
%%%%%%%%%%%%%%%
%% Section 1 %%
%%%%%%%%%%%%%%%
\section{Two State-type Model with Time-varying Effects and Time-change Covariates}
Build transition-type definition matrix on your own:
<<>>=
tra2 <- matrix(ncol = 2, nrow = 2, data = FALSE)
tra2[1, 2] <- tra2[2, 1] <- TRUE
tra2
@
Or use function \texttt{transMat} from R add-on package \texttt{mstate}:
<<>>=
library("mstate")
tra2 <- !is.na(transMat(list(c(2), c(1))))
tra2
@
Use \texttt{mplskeleton} to construct a skeleton of the model parameter list \texttt{mpl}:
<<>>=
mpl <- mplskeleton(tmat = tra2)
str(mpl)
@
Parametrize the model and complete \texttt{mpl}:
<<>>=
mpl[[1]]$bhr[[2]] <- mpl[[2]]$bhr[[1]] <- function(t){return(0.5)}
## time-varying x2 and time-varying f(x2):
mpl[[1]]$eta[[2]] <- function(x.i, t){
  ifelse(t < 5, 
         return(1.0 * x.i[1] + 0.5 * x.i[2]), 
         return(1.0 * x.i[1] + 1.0 * x.i[3]))}
## time-varying x2 and time-varying f(x1):
mpl[[2]]$eta[[1]] <- function(x.i, t){
  ifelse(t < 5, 
         return(-0.5 * x.i[1] + 0.5 * x.i[2]), 
         return( 1.0 * x.i[1] + 0.5 * x.i[3]))}
@
Simulate event histories:
<<>>=
set.seed(123)
N <- 500
X <- matrix(nrow = N, ncol = 2, rnorm(2 * N))
X <- cbind(X, X[, 2] + runif(N)/10)
colnames(X) <- c("x1", "x2.t1", "x2.t2")
head(X)
Xstruc <- matrix(ncol = 2, nrow = 2, data = 0)
rownames(Xstruc) <- c("t1", "t2")
colnames(Xstruc) <- c("x1", "x2")
Xstruc[, 1] <- 1
Xstruc[, 2] <- c(2, 3)
Xstruc
d <- simeventhistories(n = N, mpl = mpl, X = X, max.time = 10, 
                       change.times = c(5), Xstruc = Xstruc)
head(d$msm.basics)
head(d$ttsce[, 1:7])
head(d$tt.indicators)
@
Nelson-Aalen estimator of the cumulative hazard rate functions:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
plotnae(d = d$msm.basics, mpl = mpl, xlim = c(0, 15))
abline(a = 0, b = 0.5, col = "darkgrey")
@
